% Start here


\chapter{Conclusion}
\label{sec:concl}



\begin{comment}
MPD : Todo

The sequel is let as an example is this early version.

Criteria to discuss here are those which concerns all the secondary tools as open-source issues, compatibility with primary tool-chain, compatibility with eclipse,...
\end{comment}



This conclusion give a sum up of the evaluation results for each approach. The detailed results of each approach are given in the appendix.

The evaluation of Core (see Appendix \ref{sec:core}) and Why3 (see Appendix \ref{sec:why3}) have been stopped by the authors and do not appear in this conclusion.

Minus mark "-" means this criteria as not been evaluated for this approach.

Star mark "*" means this criteria has been difficult to evaluate for this approach.

The figures correspond to the sum of the scores put by one author and two assessors for each criteria and each means.
In the case of PetriNets and GnatProve, there is only one assessors thus the first figure is the sum of the 2 scores and the second in parenthesis is the result multiplied by 1.5.

The highest score is \textcolor{red}{\textbf{9}} and means that the criteria is fully respected, the lowest score is \textcolor{green}{0}.

\section{Main usage of the approach}
\label{main_usage}
This section discusses the main usage of the approach.

According to the figure \ref{fig:main_process}, for which phases do you recommend the approach (give a note from 0 to  3) :

\begin{tabular}{|l | c | c | c | c | c | c | c | c | c | c |}
\hline
&  \rotatebox{90}{GOPRR} & \rotatebox{90}{ERTMSFormalSpecs} &  \rotatebox{90}{SysML with Papyrus} &  \rotatebox{90}{SysML with EA} &  \rotatebox{90}{SCADE} &  \rotatebox{90}{EventB} &  \rotatebox{90}{Classical B} &  \rotatebox{90}{System C} & \rotatebox{90}{Petri Nets} &  \rotatebox{90}{GNATprove} \\
\hline 
System Analysis & \textcolor{blue}{5} & 1     & \textcolor{magenta}{7} & \textcolor{red}{\textbf{9}} & 3     & \textcolor{red}{\textbf{9}} & 3     & 2 & \textcolor{red}{\textbf{6(9)}}  & 2 (3) \\
\hline
Sub-system formal design  & \textcolor{red}{\textbf{9}} & \textcolor{red}{\textbf{9}} & \textcolor{blue}{6} & \textcolor{magenta}{7} & \textcolor{red}{\textbf{9}} & \textcolor{red}{\textbf{9}} & \textcolor{blue}{5} & \textcolor{blue}{5}  & \textcolor{red}{\textbf{6(9)}}   & 3 (4) \\
\hline
Software design  & \textcolor{red}{\textbf{9}} & \textcolor{green}{0} & \textcolor{blue}{6} & \textcolor{magenta}{7} & \textcolor{red}{\textbf{9}} & \textcolor{blue}{6} & \textcolor{red}{\textbf{9}} & \textcolor{red}{\textbf{9}} & \textcolor{red}{\textbf{6(9)}}   & \textcolor{red}{\textbf{6(9)}}  \\
\hline
Software code generation  & \textcolor{red}{\textbf{9}} & \textcolor{green}{0} & 3     & 3     & \textcolor{red}{\textbf{9}} & 3     & \textcolor{red}{\textbf{9}} & \textcolor{blue}{6} & 2 (3) & \textcolor{red}{\textbf{6(9)}}   \\
\hline
\end{tabular}
