\documentclass{template/openetcs_article}
\usepackage{lipsum,url}
\graphicspath{{./template/}{.}{./images/}}
\begin{document}
\frontmatter
\project{openETCS}

%Please do not change anything above this line
%============================
% The document metadata is defined below


%assign a report number here
\reportnum{OETCS/WP3a/D01}

%define your workpackage here
\wp{Work-Package 3a: ``ToolsChain''}

%set a title here
\title{openETCS Toolchain WP Description of Work}

%set a subtitle here
\subtitle{A template for short document. Adapted from report template.}

%set the date of the report here
\date{October 2012\\Revised October 2012}
\date{\today}

%define a list of authors and their affiliation here

\author{Michael Jamstram}

\affiliation{WP3a leader}
  
\author{WP3a participants}

\affiliation{OpenETCS}
    
% define the coverart
\coverart[width=350pt]{chart}

%define the type of report
\reporttype{Description of work}


\begin{abstract}
%define an abstract here
This document contents the description of work  planned for WP3a.

\end{abstract}

%=============================
%Do not change the next three lines
\maketitle
\tableofcontents
\listoffiguresandtables
\newpage
%=============================

% The actual document starts below this line
%=============================


%Start here





% Makes Marginpars easier to read
\setlength{\marginparwidth}{1in}
\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\scriptsize #1]%
{\raggedright\scriptsize #1}}

\newcommand{\oldtext}[1]{{Old: \scriptsize #1}}

\newenvironment{inoutput}
{\vspace{2mm}
\noindent
\begin{tabular}{|r|p{.7\linewidth}|l|}
\hline}
{
\hline
\end{tabular}}



This Work Package will provide the tool chain, based on formal methods, that is necessary to
design, develop, verify and validate the ETCS system. This tool chain will encompass all description levels of the system design from holistic viewpoint to code generation. Specific attention shall be paid to the semantics and the traceability of each part of the chain. The formal specification will be used further for verification and code generation.
\marginpar{(Systerel) Why to restrict the use of formal methods to verification and code generation ? Formal methods can be use at the earliest steps of the design to validate models and to  verify  traceability between steps.} The tool chain must support the following tasks:

\begin{enumerate}
\item Support the writing of the formal ETCS system
  specifications including modeling languages, graphical or
    textual editors, version management, \dots.

\item Support the writing of a formal ETCS software specifications (while a system specification describes the \emph{problem}, a software specification describes the \emph{solution}).

\item Support code generation from software specifications.

\item Support execution, debugging and simulation of software specifications.

\item Support test case/test data/test oracle generation and test procedure
execution from a software test specification.
\marginpar{(All4TEC) The software test cases should be scheduled at each step of
software development life cycle (e.g.: unit test, integration test, final
validation test)}
\marginpar{(Syterel)  If the code is automatically  generated from the software specification, is it necessary to generate software test cases e.g unit test ? }

\item Support the verification and validation of the various artefacts, including the formal ETCS specification against the textual ETCS specification.

\marginpar{MPA: should be defined, what are its relation to the system
    specification that is the requirement for the software
    specification. It should either be part of it as requirements for
    the software, or be derived from it}

\item Comply with the EN 50128 requirements to tools.

\item Support requirements tracing across all tools and build steps.

\item Analyse transitions and cooperations between models (graphical/ textual, System/ software / test,...). Several models could be successively performed from system to software level, and at the software level, at least two models could be necessary one for functional specification and one of design.

\item Provide seamless integration of the tools into one tool chain. 
\marginpar{(Systerel) "one" integrated tool chain is not necessary the best solution  to  have a certifiable approach. Indeed integration is a source of faults introduction.}
 
\end{enumerate}

\marginpar{MPA: We should state that it should allow to design the
  system/software safety case}
The tool chain definition will benefit from other R\&D projects and off-the-shelves tools. The semantics of the modelling languages shall be carefully studied.

The first goal of this WP is to identify sets of consistent languages and tools enabling the design of the system.  This will be done in close collaboration with WP2.

In order to be able to progress without depending too much on WP2 requirements and other deliverables, the subtasks shall make use of prototyping in order to gain knowledge regarding the possible modelling languages and tool platforms.  

Compliance with standards as CENELEC or 50128 could have a
significant impact in terms of workload. Moreover it should be included at the
beginning of the tools development in the Development Plan or in a Safety or
Quality Plan in close collaboration with WP4.

%=======================================================================
\section{Core Tool Chain Analyses and Recommendations}
\label{sec:core_tool}
%=======================================================================
The first task is the core tool chain analyses and recommendations. It
is concerned with the languages themselves, the tools for authoring, as well as the methods used for formalising the specifications.

The core tool will be complemented by supporting tools, as outlined in Section~\ref{sec:supporting_tools}.
Specific attention shall be paid to the semantics and the traceability of each part of the chain.


%This tool chain will encompass all description levels of the system design from holistic viewpoint to code generation.  This task is composed of the following activities:


%\begin{quotation}
%\marginpar{(mj) That's a good idea that supports Stan's idea of using prototypes.}
%(jonasHelming) For both, 1.1 as well as 1.2 it would be helpful to have a reference set of ERTMS requirements from WP2. This should be small sub-set of the specification, but be somehow representative. Using this sub-set would ake it much easier to compare different languages and tools. It can be used as a test set. I think this might be even more useful then detailed requirements for the toolchain, as we have quite some compentence on that in WP3a I think. Of course we will work on the requirements from WP2, but to get startedt with a comparison of existing tools, I think a test set might be the most required piece. Any opinions on that? 
%\end{quotation}

%-----------------------------------------------------------------------
\subsection{Identify and define the potential modelling languages}
\label{sec:language}
%-----------------------------------------------------------------------

The objective of this subtask is the identification of the modeling language (or multiple languages), based on the analysis from WP2, by prototyping.  For each candidate, a small subset of the ERTMS specification will be modelled.  The languages may have to be adapted in the process.
%\marginpar{(ALL4TEC) ALL4TEC is candidate to contribute for modeling on the basis of Subset 026 for example …}
The completed prototypes are subsequently evaluated against the requirements from WP2.

This work is based on merit: If a suitable language is identified, but no partner steps up to model the prototype, it will simply not be considered.

%\marginpar{(mj) I think there is a huge overlap with WP2.  Their findings may reduce the possible target languages drastically, and by their requirement must take everything written here into account anyway.}
%\marginpar{(SP) I agree with michael, so I removed the language analysis, and replaced it by prototyping using the candidate languages selected by WP2.}
%WP2 shall identify the candidate languages suitable to model the Subset-026. 

%The following prototypes shall be developed, one for each candidate language outlined by WP2 State of the Art Analysis, and this for a representative subset of Subset-026 requirements. 

%\marginpar{(mj) I would suggest a dedicated sesction to prototyping.  Keep in mind that prototyping consists of at least two parts: (1) modeling the reference requirements; (2) building a prototypical tool platform}

Given that different levels of abstraction have to be addressed in the
design phase, several languages may be necessary to handle all design
phases. Here, the semantics of the languages is the key point; it
shall be accurately adapted to the level of description required in
each phase of the design.

The identification and definition will distinguish between
%\marginpar{(jp) Added distinction between wide-spectrum formalisms (e.~g., UML, SysML, B) and DSLs}
\begin{itemize}
\item wide-spectrum modelling languages (e.~g., UML, SysML, B,...) suitable for a wide variety of modelling domains,
\item domain-specific languages (DSL) designed and optimised 
for application in a specific application domain 
only.
\end{itemize}
For wide-spectrum languages their metamodels\footnote{Recall that metamodels specify the syntax
and static semantics of a modelling formalism.} will be analysed with respect to their expressive power
and resulting adequateness for designing ERTMS models.
For DSL candidates the associated meta-metamodels\footnote{Recall that the meta-metamodel 
specifies the capabilities to define language elements and their static semantics in a DSL.} 
will be analysed with respect to their
capabilities to support language extensions that may become necessary for novel releases of the ERTMS 
specification the in the future. 

\begin{inoutput}
Input: & WP2: List of suitable languages (based on State of the Art Analysis) & Oct-12 \\
Input: & WP2: Small subset of ERTMS requirements that is representative & Oct-12 \\
Input: & Those WP2 Requirements that are sufficient to evaluate a target language & Jan-13 \\
\hline
Output: & Formal Model representing the sample spec, one for each candidate & Feb-13 \\
Output: & Documentation of the changes to each language used (if any) & Feb-13 \\
Output: & Evaluation of the models against the WP2 requirements & Mar-13 \\
Output: & Decision on the final language choice(s) & Apr-13 \\
\end{inoutput}
\marginpar{(Systerel) A template of criteria to select the language and tools should be defined.}

%-----------------------------------------------------------------------
\subsection{Identify and compare existing tools}
\label{sec:tool}
%-----------------------------------------------------------------------

Corresponding to Section~\ref{sec:language}, the objective of this subtask is the identification of the target tool, based on the analysis from WP2, by using it for the prototyping described in Section~\ref{sec:language}.

The experience with the tools will be recorded, and the tools will be evaluated against the requirements from WP2.
%\begin{quotation}
%\marginpar{(mj) I'd assume that's part of WP2's ``state of the art''.  I'll make it explicit in the input.}
%(jonasHelming) I think we had this task in the FPP before, was it removed? I think that is already a task we could get started with. I think there was already a proposal on the mailing list to create wiki pages for the tools suggested by the partners. It would make sense, tough to prepare a template for this.
%\end{quotation}

\begin{inoutput}
Input: & WP2: List of suitable tools (based on State of the Art Analysis) & Oct-12 \\
Input: & Those WP2 Requirements that are sufficient to evaluate the tool & Jan-13 \\
\hline
Output: & Experience report for each candidate tool & Feb-13 \\
Output: & Documentation of the changes to each tool (if any) & Feb-13 \\
Output: & Evaluation of the tools against the WP2 requirements & Mar-13 \\
Output: & Decision on the final tool choice(s) & Apr-13 \\
\end{inoutput}

%-----------------------------------------------------------------------
\subsection{Identify the tool platform}
\label{sec:tool_platform}
%-----------------------------------------------------------------------

There is a distinction between tool (Section~\ref{sec:tool}) and tool platform: The tool is the core that processes the language, and typically also has an editor.  The tool platform is language independent, but provides mechanisms to integrate various tools.  For example, Eclipse is a tool platform.  The Java Development Tools (JDT) are an extension to Eclipse that allows working with the Java programming language.

As the toolchain will consist of many tools that must work together seamlessly, it should be analysed independently from the tool.  A tool will typically suggest a certain tool platform.  The aim of this task is the identification of a tool platform for each candidate tool from Section~\ref{sec:tool}.

\marginpar{(ALL4TEC) an interface facility should be examined to facilitate the data exchange between the tools constituting the tool chain (e.g.: XML definition) }
%These tools enable designers to model using the languages defined before. The interoperability of the tools is a key point to address traceability of the different models in the design process. Tools may already exist and may be used “off-the-shelves” or may be developed in the WP. As example, if a standard modelling language is selected such as UML/SysML, open tools already exists (Topcased); if a specific modelling language is needed, WP actors have to develop a grammar or meta model to develop such particular modelling tool.

\paragraph{irit-laas} we could stress that there are several levels
of interoperability that can be supported by the toolchain. Taking the
example of Eclipse, the most basic level provides support for plugins
(OSGi); that is the ability to have several tools inside a common
platform. In the project, we will also need support for
interoperability at the data level, which can be found in the Eclipse
world with the Modeling Framework; that is the ability to have tools
that converse/interact using a standardized (low-level)
syntax. Finally, we could also require interoperability at a
``semantical'' level---for example with support for linking objects in
different modeling languages---or ask for more basic services, such as
serialization and versioning. We believe that, for the project, we at
least need interoperability at the data level.

\begin{inoutput}
Input: & List of target platforms, based on the tools being evaluated (\ref{sec:tool}) & Oct-12 \\
& Those WP2 Requirements that are sufficient to evaluate a target language & Dec-12 \\
\hline
Output: & Evaluation of each tool platform against WP2 requirements, independent of target tool & Feb-13 \\
Output: & Evaluation of tool platform in the context of specific target tools & Mar-13 \\
Output: & Selection of Tool Platform (and reasoning) & Apr-13 \\
\end{inoutput}

%-----------------------------------------------------------------------
\subsection{Identify Development Method, including Elicitation Techniques}
%-----------------------------------------------------------------------

Analyse requirement elicitation techniques in order to define a strategy to derive OpenETCS formal model requirements from ERTMS SRS

Without a suitable method, most tools have only limited use.  Formalising a specification of the size of ETCS requires a suitable method.  The method is probably relatively independent from the language used.

The objective of this subtask is the identification of suitable methods and their evaluation while prototyping.

\marginpar{(SP) I would suggest to remove that subtask.
- This belongs to WP2.
- Requirement elicitation technique is not a tool, and moreover
- ERTMS SRS (Subset-026) already include the requirements.
}
\paragraph{Christophe GASTON}
A classical modelling process starts by defining high level models of systems that shall be refined step by steps in order
to make implementation choices. From a syntactical point of view, the model transformation techniques described below are good technological candidates. However, at the semantical level, refining a model into a more concrete one requires first to define a refinement correctness relation in order to ensure properties preservation. Moreover, systems that will be considered in the project will be based on concepts of interacting processes (since those systems are distributed). Therefore we need to identify techniques allowing us to start from system level properties ({\em i.e.} specifying global behaviors) and to deduce what properties each process should satisfy ({\em i.e.} what behavior should have each process) so that the global cooperation of all processes realize the system properties. In this subtask we will identify potential solutions to that problem.
\marginpar{(ALL4TEC) Christophe Gaston comments address several previous chapters of the DOW. It could be useful to re dispatch them. }

\paragraph{irit-laas} we are also in favor of removing this
subtask. One drawback of mandating a specific method for developing
the models (e.g. refinement or compositional) is that you could rule
out some of the tools identified in Sect. 1.2.

More generally, we believe (based on our previous experiences in other
domains) that we will end up with several formalisms, targeting
different aspect of the specification. We think that we should keep in
mind an approach based on modeling different aspects of the problem
with languages supporting different viewpoints and that could work at
different level of abstractions. With this kind of
approach---concerning the verification of the final artifacts---we
will also need to check the compatibility of the semantics of the
modeling languages that address overlapping viewpoints. There are two
problems here. First, when dealing with an heterogeneous
specification, we need a common semantical basis to check the
compatibility of the models. More pragmatically, when we deal with two
models (expressed in a different language) that describe the same part
of the system, we need to show that they are consistent with each
other.

\marginpar{(Systerel) Requirements elicitation is the first step of the process, but we can decide that requirements are already elicitate in the subset-026. However, how to manage the requirements traceabilitty and how to deal with traceability between models and activity is a main challenge in developping critical and complex systems. Thus, we think it is important to define how to support requirement traceability.}

\begin{inoutput}
Input: & WP2: List of suitable methods (based on State of the Art Analysis) & Oct-12 \\
Input: & Those WP2 Requirements that are sufficient to evaluate a target method & Jan-13 \\
\hline
Output: & Experience Report on applying the method while prototyping (\ref{sec:language}) & Feb-13 \\
Output: & Evaluation of the methods against the WP2 requirements & Mar-13 \\
Output: & Decision on the final method & Apr-13 \\
Output: & Documentation of adapted method & ongoing \\
\end{inoutput}

%=======================================================================
\section{Supporting Tools Analyses and Recommendations}
\label{sec:supporting_tools}
%=======================================================================

The languages and tools  of the core chain have to be complemented to support a number of activities that are crucial for the project. The purpose of this task is to compose a list of tools which may be
used in performing or supporting  activities, analyse their
suitability and propose a selection for inclusion in the openETCS tool
basis. This has to be coordinated with 
\begin{itemize}
\item the core tool chain selection (suitable  tools should be
  available), 
\item the definition of the overall development process, as that defines the steps to be performed and 
\item the safety analyses to  conclude if the proposed tools are suitable to  respond to standard criteria for certifiable tool chain.
\end{itemize}
  The flow of information in the coordination is bidirectional--e.g., the process
steps should ideally be taylored towards realizability with good
support by available tools. Techniques complementing the
tool-supported steps must be included in the process description. 
  
The following activites have to be take in account :
\begin{itemize}
\item Model Transformation and code generation
\item Validation and verification
\item Requirement management
\item Safety analyses Tools
\end{itemize}


%-----------------------------------------------------------------------
\subsection{Model Transformation and Code Generation}
%-----------------------------------------------------------------------

Analyse model transformation techniques and tools in order to refine the specification from one description level to another.

%\paragraph{Stanislas Pinte}
%In our product ERTMSFormalSpecs (http://www.ertmssolutions.com/ertms-formalspecs/) we use a single, unified model, that is supporting 
%complete Subset-026 logic, with full traceability. 
%If that approach works, why do we need several specification levels? 
%\paragraph{irit-laas} we propose to merge sections 2.2 and 2.3 since
%they are basically the same kind of activity. It could have the added
%benefit to balance the ``size'' with respect to the verification
%activities in Sect. 2.1


Analyse the code generation strategy.



%Additionally, prototypes should be developed for each candidate modelling language, and for each candidate target language. 
%\paragraph{C.~Braunstein, J.~Peleska (Uni Bremen)}
%The code generation should be compatible with a domain framework including, for
%example, the designated operating system API to be applied in openETCS. This
%should be decide within the tool chain. \\
%Another issue is that code generation should be covered down to binary code, so that V\&V arguments
%may be applied down to the level of binary code.
%
%A more detailed view of these issues may be find in the PositionPaper directory
%(Uni\_Bremen\_proposed\_Workflow.pdf).\\

\begin{inoutput}
Input: & WP2 Requirements & Jan-13 \\
\hline
Input: & WP3-a  Decision on the final language choice & Apr-13 \\
\hline
Output: & Model Transformation Strategy & Sept-13 \\
\hline
Output: & Code Generation Strategy & Sept-13 \\
\end{inoutput}


%-----------------------------------------------------------------------
\subsection{Validation and Verification Tools}
%-----------------------------------------------------------------------

In the development of a safety-critical rail system like the ETCS OBU,
several kinds of verification and validation activities have to be
performed. One the one hand, the relevant standards require the
verification of the \emph{safety aspect} of every major design
step. They also list constraints on methods (and tools) which may be used for these
purposes.  On the other hand, any viable development process for a
complex real-time system like the OBU needs early validation of design
artifacts. This concerns for instance functionality and real-time
properties beyond the safety-related issues. The formalisation of
design artifacts which comes with a model-based process offers the
possibility to employ tools to a substantial degree for these
activities. 


 
% Identify potential verification tools with regard to modelling techniques; verification techniques shall be investigated.

%\paragraph{Hardi Hungar}
%Verification tools resp. techniques are very important in the development of a safety-critical system like the ETCS OBU. According to the relevant standards (most prominently the EN 50128 of the CENELEC family), every design step has to be verified. Assuming that models will constitute artifacts of the development – and are not just used for explanatory purposes – it is necessary to be able to establish the correctness of each refinement step. Or, to put it differently, the tool chain needs a concept for seamless verification, preferably tool-based, to be fit for its purpose.

Validation and Verification activities have different goals and methods :
\begin{description}
\item [Validation] aims to prove that the build software fulfills the user needs, in particular with respect to safety and quality requirements.
\item [Verification] aims to prove that the build software fulfills the requirements of that
phase with respect to completeness, correctness and consistency.
\end{description}



The choice of VnV methods depends on the kinds of properties which are treated.

%\marginpar{(ALLTEC) Notice: CENELEC verification activities cannot be systematically addressed with test and tool. For example, source code review could be required to check safety properties of code.}
%\paragraph{Stanislas Pinte}
%I agree with Hardi.
%In my opinion, the model should include the tests of the model, so that it could be verifiable in a "model-in-the-loop" fashion.
%It doesn't have to be model proving (that I think belongs more to other WPs) rather model testing.%
%Inside our http://www.ertmssolutions.com/ertms-formalspecs/ approach, we assume the following:

%- Model tests are part of the model

%- Model should be 100\% covered by tests (proved by model coverage reports)

%- Toolchain must support developing, executing and debugging tests

%Model verification also includes:

%- verifying that the model corresponds to the original requirements specifications (in our case, UNISIG Subset-026 BL3). ERTMSFormalSpecs also supports marking model artifacts as "verified" against source requirements.

%- verifying that 100\% of source requirements are traced against one or more model artifacts (proven by traceability reports)

%I would think that such verifications are indeed in the critical path...i.e. if not implemented we shall not be able to have a fully functional model.

%\paragraph{All4TEC}
%ALL4TEC has already experienced tools at several levels of critical railways system and software developments. 
%In order to verify the work, tools like CORE (Vitech) could be used for system technical specifications elaboration (it is a kind of SysML approach leading to definition of the system functional and physical architectures).
%In on other way Safety Architect could be used for safety studies on various
%models at diverse levels of definition (system, software, functional, organic
%etc.).\\
%The following tool chains have been experienced for software testing: 
%For Unitary Test (UT): MATLAB SIMULINK / DESIGN VERIFIER
%For Validation: MATELO /TESTAND /VERISTAND


\subsubsection{Functional properties}
%.......................................................................

First, we have to show that all of the components of the system behave
\marginpar{(DLR) Some further rewriting of this subsection
  and the following is needed in order to get a consistent
  description. Concerning the current text, one could broaden the
  scope of functional verification, including early design steps.) }
as they are intended. This includes at least proving that low-level requirements meet
high-level requirements, and in turn that the code meet these low-level
specifications. 

One possibility is to use a completely certified toolchain up
to code generation, some formal methods have shown their effectiveness in many industrial cases. Failing that ({\it i.e.} if not all the code is generated
and/or if the generation cannot be trusted), we will need to have formal
specifications and to verify the code against them. Hoare logic-based tools
seem like a good approach in this case, as well as of course test cases
generation according to a suitable coverage criterion.

Secondly, we have to  show that the high level requirements have been completely specified. This is usually  achieved by coverage technics. 

\subsubsection{Non Functional properties}
%.......................................................................

Such properties include all aspects that are related to the nominal behavior
of the components. In particular, this concerns the following points:
\begin{itemize}
\item schedulability and Worst Case Execution Time (WCET)
\item data dependencies between outputs and inputs
\item modelling and process requirements for SIL4 systems
\end{itemize}

Schedulability can be done on the models, based on hypotheses for the WCET of
single tasks, but verifying these hypotheses require a full knowledge of
the code, the underlying hardware, and the compilation toolchain.

Data dependencies can also be assessed on the models, but as in the previous
subsection, some verification on the code itself might be required. Static
analyzers should be able to handle that.

Traceability  items can help  to  check  that proces requirements are satisfied, code reviews allowed to  verify modelling requirements.

%.......................................................................
\subsubsection{Safety properties}
%.......................................................................

\marginpar{(Systerel) a section Safety analysis Tools has been added : this subject is very important for safety critical systems and not yet very well supported.}

This must ensure that components are always able to perform their work.
It includes in particular:
\begin{itemize}
  \item Absence of runtime error (again if the code generation does
        not guarantee it by construction. Static analysis can be employed
        there also).
  \item Fault tolerance against unexpected input (safety analyses can raise this unexpected input).
  \item High-level safety requirements (formal raisonning, test or simulation can show safety requirements are satified).
\end{itemize}

\begin{inoutput}
Input: & WP2 Requirements & Jan-13 \\
Input: & Those sufficient elements of WP4 VnV Plan to evaluate tools & Jan-13 \\
\hline
Output: & VnV tool choice & Apr-13 \\
\end{inoutput}



%-----------------------------------------------------------------------
\subsection{Test Automation Tools}
%-----------------------------------------------------------------------

Testing 
\marginpar{(jp) Some standards (e.~g.~RTCA DO178B/C) regard testing as a part of the verification
process. I think it is clearer, however, to distinguish between verification in the sense of
static analysis, abstract interpretation, formal verification, WCET analysis and testing in the sense
of dynamic testing in an explicit way in this DoW}
\marginpar{(Systerel) in railway standard (e.g. EN50128), testing and verification are linked. However  it is usual  in railway industry  to  replace unit tests for instance by formal verification and automatic code verification for functionnal  parts. Moreover integration of HW/SW is out of the scope of the OpenETCS projets (because due to industiral choices), thus we do not think this new section is necessary.} 
is mandatory for the development of safety-critical systems, because no system is allowed
to become operative without experimental evidence.
In all situations where formal verification cannot be exhaustive, because 
 the size and complexity of the software or integrated HW/SW system exceeds the
capabilities of the formal verification power available, testing or simulation is applied 
to achieve at least partial verification for the cases considered.

In the openETCS context, testing is required in two areas.
\begin{itemize}
\item Software and HW/SW integration testing: the overall test objective is to
investigate the correct cooperation between several SW and HW components.
\item Component (unit, module,  thread, process) testing to investigate (1) their functional
correctness 
and (2) the consistency between models and
 object code generated from the models. The former use case is mandatory if the model has not already been exhaustively verified. The latter use case is relevant when utilising code generators 
 that have not been validated, so that we cannot rely on the correctness of the model-to-code transformation.
 \marginpar{(Systerel)  in the case of code generator, testing can be replace by the use of a second generator and some process verification.}
\end{itemize}
In any case a high degree of automation is desirable to support an effective testing process.
Available test tools will be analysed with respect to the following properties.
\begin{itemize}
\item Availability in open source
\item Suitability for the preferred tool platform 
\item Support for different test levels (unit, integration, HW/SW integration, hardware-in-the-loop system testing)
\item Support for automated test procedure / test suite execution
\item Support for automated test case / test data / test oracle / test procedure generation from
the preferred type of models selected in WP2 / WP3a. 
\item Support for automated tracing from requirements to test cases to test procedures to test results and back
\end{itemize}



\begin{inoutput}
Input: & WP2 Requirements & ??? \\
\hline
Output: & Test automation tool choice with experience report & ??? \\
\end{inoutput}


%%-----------------------------------------------------------------------
%\subsection{Schedulability}
%%-----------------------------------------------------------------------
%
%\marginpar{CEA: this section could disappear and the content would be dispatched in subsection "`non functional properties" of Section "`Verification"}
%\marginpar{(ALL4TEC)
%We agree with the remark on schedulability (e.g.: should be removed).
%}
%\marginpar{irit-laas: Should we add something about scheduling
%  ``synthesis'' to the subsection on schedulability? Maybe the
%  scheduling method/startegy is mandated by the SRS? Same remark for
%  adding WCET analysis to section 2.4.}
% 
%Analyse schedulability tools.
%
%\begin{inoutput}
%Input: & WP2 Requirements & ??? \\
%\hline
%Output: & Schedulability Strategy & ??? \\
%\end{inoutput}

%-----------------------------------------------------------------------
\subsection{Capture Additional Requirements}
%-----------------------------------------------------------------------

Capture wishes/requirements on how to support the designer in their activities.

\begin{inoutput}
Input: & Designer Wishes and Requirements & ongoing \\
\hline
Output: & Captured and organised designer requirements & ??? \\
\end{inoutput}


%-----------------------------------------------------------------------
\subsection{Safety analyses Tools}
%-----------------------------------------------------------------------


The purpose of this subtask is to propose a list of tools to  support  safety  analyses activities. this task  is to  coordinate with  the requirements issues of WP2 and the needs of WP4 in regards of safety cases definition.

Examples of interresting tools deals with fault-trees and FMEA generation. 
  
 


\begin{inoutput}
Input: & WP2 Requirements & Jan-13 \\
Input: & Those sufficient elements of WP4 Safety case to evaluate tools & Jan-13 \\
\hline
Output: & safety analyses tools choices & Apr-13 \\
\end{inoutput}



%=======================================================================
\section{Define and Develop Tool Chain}\label{sec:devtoolchain}
%=======================================================================

\marginpar{(mj) I see "make" not as an option.  Considering the resources available for this project, we will tailor (and extend) an existing tool platform.}

This 
\marginpar{(jp) Please review this section with me and check it w.r.t.~completeness}
subtask defines and develops the tool chain and the infrastructure enabling its evolution and maintenance. First of all, a "make or reuse" decision about the components of the tool chain has to be made. Then a common development infrastructure has to be defined or chosen in order to integrate all the tools (Eclipse like infrastructure). Finally, the subtask achieves the development and the integration of the tools.

%-----------------------------------------------------------------------
\subsection{Overall Tool Architecture}
%-----------------------------------------------------------------------

Once language, tools and tool platform have been identified, the architecture can be defined using that as the foundation.  The architecture also contains the specification of tool interaction mechanisms (model changes might trigger code generators, model or code changes might trigger regression tests etc.) 

Note that this may not be that much work:  A tool platform like Eclipse essentially defines the overall architecture already.  Further, using an agile approach, it is perfectly acceptable if the system changes over time (i.e. APIs change, etc.), as long as the proper mechanisms are in place, like automated testing.

%-----------------------------------------------------------------------
\subsection{Development Infrastructure}
%-----------------------------------------------------------------------

To allow robust distributed development, care must be taken in setting up a functioning infrastructure.  This includes 
\begin{itemize}
\item a continuous automated build system, 
\item mechanisms to upgrade tools in the platform, 
\item mechanisms to add tools to the chain at a later stage,
\item tool chain documentation system. 
\end{itemize}
 The effort for this must not be underestimated.

%-----------------------------------------------------------------------
\subsection{Decomposition and Distribution of work}
%-----------------------------------------------------------------------

Another major task is the robust decomposition of the tool chain and its 
distribution and tracking of the various components.  
Specifically, robust integration tests and version management for the tool chain are crucial\footnote{The related activities are often called {\it tool qualification}.}.



\paragraph{Inputs and Outputs for WP3a, Section~\ref{sec:devtoolchain}.}{~}\newline


\begin{inoutput}
Input: & WP3a, Section~\ref{sec:core_tool}: Formal Model representing the sample spec, one for each candidate  & ??? \\
Input: & WP3a, Section~\ref{sec:core_tool}: Decision on the final language choice(s) & ??? \\
Input: & WP3a, Section~\ref{sec:core_tool}: Experience report for each candidate tool in the core tool chain& ??? \\
Input: & WP3a, Section~\ref{sec:core_tool}:Decision on the final tool choice(s)  & ??? \\
Input: & WP3a, Section~\ref{sec:core_tool}: Selection of Tool Platform (and reasoning)  & ??? \\
Input: & WP3a, Section~\ref{sec:core_tool}: Decision on the final (development) method  & ??? \\
Input: & WP3a, Section~\ref{sec:supporting_tools}: Verification and test tool choice(s)  & ??? \\
Input: & WP3a, Section~\ref{sec:core_tool}: Captured and organised designer requirements  & ??? \\
%Input: &   & ??? \\
%Input: &   & ??? \\
%Input: &   & ??? \\
\hline
Output: & Specification of tool interoperability mechanisms & ??? \\
Output: & Specification of core and support tool chain architecture and its embedding into the platform & ??? \\
Output: & Infrastructure evolution strategy & ??? \\
Output: & Tool chain qualification test suite & ??? \\
\end{inoutput}



%=======================================================================
\section{Develop Open Source Ecosystem}
%=======================================================================
The goal of this task is the development of an open-source ecosystem for the toolchain under development and all its components as well as for the implementation of the ETCS system. This ecosystem defines the license model to be used, the project infrastructure, usage and contributions of existing open-source projects and the process to coordinate the development efforts from different partners. Additionally as part of this task, suggestions and guidelines for the projects infrastructure are developed. The goal of the ecosystem is to facilitate the collaboration of the industrial partners and enable long-term maintenance of the outcome of the development. For the ecosystem, well-established open source ecosystems, such as the Eclipse ecosystem shall be used as templates.

As the development of an ecosystem needs to react on the on-going project and adapt to needs, the ecosystem has to evolve over the project duration. As a first output, this taks will produce initial proposals for the license model, the process, as well as the infrastructure. Subsequently, these proposals can be adapted based on the feedback from all project partners.

As the license model and the open source process have high impact on the overall project, the task will only propose solutions. These solutions are based on requirements given by WP1. Final decisions in this area have to be made by WP1. 

\begin{inoutput}
Input: & Feedback from project partners (All WPS) & continuosly \\
Input: & WP1: Requirements for the open source ecosystm & continuosly \\
Input: & WP1: Legal decisions & continuosly \\
Output: & Initial proposal for license model & Aug-30 \\
Output: & Initial proposal for open source development process & Aug-30 \\
Output: & Initial proposal for infrastructure and tools & Aug-30 \\
Output: & Adaptation of license model & continuosly \\
Output: & Adaptation of open source development process & continuosly \\
Output: & Adaptation of infrastructure and tools & continuosly \\
\end{inoutput}


\end{document}
Local Variables:
ispell-local-dictionary: "english"
End:
